A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-11-18 tex/codatadp.tex
more

  * briefly: datatypes and why they are useful

@@

  * single decision procedure for datatypes and codatatypes
  * datatypes are implementation following the lines of Barrett et al.
  * codatatypes were added later, motivated by CVC4 as Isabelle backend

@@

  * codatatypes: from a theoretical and implementational point of view, like
      datatypes but:
    * codatatypes are never empty (e.g. finite streams are rejected)
    * no acyclicity (e.g. xs = lcons(0, xs) is satisfiable)
    * instead: bisimilarity

@@

    * point to Barrett et al. for SMT datatypes
    * additional ones since then (e.g. strings?)
    * anything about codatatypes?
      * proof assistants like Agda, Coq, etc. have them
      * Dafny, CoALP
      * also a lot of theoretical research, some of which is loosely connected,
        e.g. decision procedure for corecursive functions (Henning in Nijmegen)

@@

>>>>

  * briefly: datatypes and why they are useful
    (terminology: freely-generated, inductive, algebraic, ..., sometimes
    with different meanings; we'll clarify below)
    * Hoare's recursive data structures
    * comparatively easy to reason about and automate
    * common in practice, esp. in computer science applications

@@

  * single decision procedure for datatypes and codatatypes
  * datatypes are implementation following the lines of Barrett et al.
  * codatatypes were added later, motivated by the use of SMT solvers as
    backends to proof assistants (more specifically, CVC4 to Isabelle/HOL)

@@

  * codatatypes: from a theoretical and implementational point of view, like
      datatypes but:
    * codatatypes are never empty (e.g. finite streams are rejected)
    * no acyclicity (e.g. xs = lcons(0, xs) is satisfiable)
    * instead: bisimilarity
    * (also: ``enumeration'', e.g. of streams)

@@

    * point to Barrett et al. for SMT datatypes
      * about their own work, they say: "our focus is on generality and
        efficiency rather than immediacy of implementation"
      * (deal more directly with finite sorts than Barrett et al., Section 6.1)
    * additional ones since then (e.g. strings?)
    * anything about codatatypes?
      * proof assistants like Agda, Coq, etc. have them
      * Dafny, CoALP
      * also a lot of theoretical research, some of which is loosely connected,
        e.g. decision procedure for corecursive functions (Henning in Nijmegen)

    * ODDITY: Oppen 1980: single-constructor, recursive -- infinite values?

@@

  * how to deal with "wrong" selectors, e.g.
      hd(nil1) = hd(nil2)?
  * ODDITY: Barrett et al. overspecify things (not good w.r.t. SMT-LIB 2, or
    other solvers, e.g., Z3)
